### 2014-09-09 writeups/context-dependence.tex
tree inference working IF columns are in the right order

  \item Compute $M_2(v,z) = \sum_{v \in \calS^{m+2\ell}}  p_{m+2\ell}(t_{AB},u,v) \pi(u) M_1(u,z)$ (still a $|\calS|^{m+2\ell} \times |\calS|^m$ matrix)

@@

Fix one tip of the tree, $a$, to be the taxa where we count ``long'' sequences, let $\rho$ be the root, and let $b_1, \ldots, b_n$ be the remaining tips,
where we observe ``short'' sequences.
Let $v_0=\rho, v_1, \ldots, v_\ell = a$ be the path from the root down to $a$,
let $u_1, u_2, \ldots, u_m$ be the remaining internal nodes,
and for each $v_i$ let $u(v_i)$ be the offspring of $v_i$ that is not one of the other $v$.
For each node in the tree,
we will compute the matrix of probabilities that a ``long'' sequence observed at that node
is matched by the ``short'' sequences seen on all the tips of the tree
further away from $a$ (i.e.\ below that node if it was rooted at $a$).
For a node $w$ this will be stored as a $|\calS|^{m+2\ell} \times |\calS|^{cm}$ matrix, for some $c$,
and denoted $M(w)$.

@@

  \item For each $1 \le k < \ell$,
    \begin{enumerate}
      \item let
        \begin{align}
          M(v_k)_{x,(y,z)} = \left( e^{t(v_k) G} M(v_{k-1}) \right)_{x,y} \left( e^{t(u(v_k)) G} M(u(v_{k-1})) \right)_{x,z}  .
        \end{align}
    \end{enumerate}

>>>>

  \item Compute $M_2(v,z) = \sum_{u \in \calS^{m+2\ell}}  p_{m+2\ell}(t_{AB},u,v) \pi(u) M_1(u,z)$ (still a $|\calS|^{m+2\ell} \times |\calS|^m$ matrix)

@@

Fix one tip of the tree, $a$, to be the taxa where we count ``long'' sequences, let $\rho$ be the root, and let $b_1, \ldots, b_n$ be the remaining tips,
where we observe ``short'' sequences.
Let $v_0=\rho, v_1, \ldots, v_\ell = a$ be the path from the root down to $a$,
let $u_1, u_2, \ldots, u_m$ be the remaining internal nodes,
and for each $v_i$ let $u(v_i)$ be the offspring of $v_i$ that is not one of the other $v$.
For each node in the tree,
we will compute a matrix of probabilities:
for nodes not on the path from $a$ to the root,
we compute the chance that that each ``long'' sequence observed at that node
is matched by each possible combination of ``short'' sequences on the tips of the tree below it.
For nodes on the path from $a$ to the root, we compute the probability of observing
each ``long'' sequence at that node, along with the combinations of ``short'' sequences
on all tips further away from $a$ (i.e.\ below that node if it was rooted at $a$).
For a node $w$ this will be stored as a $|\calS|^{m+2\ell} \times |\calS|^{cm}$ matrix, for some $c$,
and denoted $M(w)$.
Rows of $M(u_j)$ sum to 1,
while each entire matrix $M(v_i)$ sums to 1;
the difference being that $M(v_i)$ sums over the prior at the root,
and $M(u_j)$ takes the value at $u_j$ as given.

@@

  \item For each $1 \le k < \ell$,
    \begin{enumerate}
      \item let
        \begin{align}
          M(v_k)_{x,(y,z)} = \left( e^{t(v_k) G} M(v_{k-1}) \right)_{y,x} \left( e^{t(u(v_k)) G} M(u(v_{k-1})) \right)_{x,z}  .
        \end{align}
    \end{enumerate}
